A Data-Driven Framework for Neural Field Modelling

### 2010-11-18 ltx/first_order_framework/LaTeX_Manuscript.tex
made changes from Mikes comments

The mapping between the membrane voltage and the electrophysiological data, denoted by $\mathbf y_t$, is modelled using the observation function that incorporates sensors with a spatial extent by
\begin{equation}
    \label{eq:ObservationEquation}
	\mathbf{y}_t =
	\int_{\Omega}{
	    m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
	\, d\mathbf{r}'} + 
	\boldsymbol{\varepsilon}_t, 
\end{equation}
where $m\left(\mathbf{r}_n-\mathbf{r}'\right)$ is the observation kernel, $\mathbf{r}_n$ defines the location of the electrodes in the field, $n=1,...,N$ indexes the electrodes and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$ denotes a multivariate normal distribution with mean zero and covariance matrix $\boldsymbol{\Sigma}_{\varepsilon}$.

@@

where $\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}$ denotes the $i^{th}$ column of $\boldsymbol{\Psi}(\mathbf{r}')$ which is a $L \times n_{\theta}$ matrix, where $L$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity kernel basis functions. This matrix is defined analytically (see the supplementary (\dean{Appendix}) material for analytic convolution of two Gaussians), since it is constant with respect to the states and the unknown parameters. Substituting equation~\ref{eq:DefPsi} into equation~\ref{eq:ReducedForm} gives

@@

Now we define the state disturbance as
\begin{equation}\label{eq:Wt} 
	\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} ) \, d\mathbf{r}},
\end{equation}
which is a zero mean normally distributed white noise process with covariance (see the supplementary material \dean{(Appendix)} for the derivation)
\begin{equation}
	\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top} \, d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
\end{equation}
The observation equation of the reduced model is found by substituting equation~\ref{DefFieldDecomp} into equation~\ref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_t\, d\mathbf{r}'} + \boldsymbol{\varepsilon}_t. 
\end{equation}
The observation equation is linear and can be written in the more compact form
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is 
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}') \, d\mathbf{r}'.
\end{equation}
% \begin{equation}
% 	\mathbf{C} = \left[
% 	\begin{array}{{ccc}} 
% 		c_{1,1} & \dots & c_{1,L} \\
% 		\vdots & \ddots & \vdots \\
% 		c_{N,1} & \dots & c_{N,L} 
% 	\end{array}
% 	\right] 
% \end{equation}
% and 
% \begin{equation}
% 	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}') \, d\mathbf{r}'. 
% \end{equation}
Now we have the final form of the state-space model where
\begin{equation}\label{eq:finalformstatespacemodel}
	\mathbf{x}_{t+1} = Q(\mathbf{x}_t) +\mathbf{e}_t
\end{equation}
\begin{equation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t
\end{equation}
and 
\begin{equation}\label{eq:QmatrixForSigmapoints}
	Q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \, d\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t.
\end{equation}

@@

Let the spectral representation of the post-synaptic membrane voltage field at time $t$ \parham{and at spatial frequency $\boldsymbol\nu$} be denoted by $V_t(\boldsymbol{\nu})$. According to Shannon's sampling theorem, the field must be spatially band-limited for an accurate reconstruction using spatially discrete observations. Nevertheless, an approximate reconstruction can be obtained if the field is approximately band-limited with 
\begin{equation}
	V_t(\boldsymbol{\nu}) \approx 0 ~ \ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c,
\end{equation}
where $\boldsymbol{\nu}_c$ is a cutoff frequency (typically taken as the -3~dB point). Given such a band-limited field, the distance between adjacent sensors, $\Delta_y$, must satisfy 
\begin{equation}
	\label{eq:MinimumSensorDistance} \Delta_y \leq \frac{1}{2\rho_y\boldsymbol{\nu}_{c}}, 
\end{equation}
where $\rho_y \in \mathbb{R} \ge 1$ is an oversampling parameter. This condition must be satisfied to avoid spatial spectral aliasing effects when reconstructing the hidden dynamic field, $v_t(\mathbf{r})$, using the sampled observations, $\mathbf{y}_t$.

In practice, it is difficult to estimate the bandwidth of the cortex using traditional electrophysiological measurements, possibly preventing placement of sensors in accordance to equation~\ref{eq:MinimumSensorDistance}. However, we envisage it may be possible to estimate the spatial bandwidth using other modalities with higher spatial resolution such as fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Nevertheless, spectral aliasing can still be avoided by a proper choice of the spatial sampling distance given the sensors' spectral characteristics. The spatial extent of the sensors results in a spectral low-pass action, thus providing spatial anti-aliasing filtering. Such sensors limit the spatial bandwidth of the observation to a value  denoted by $\boldsymbol{\nu}_{cy}$. Aliasing can then be avoided by positioning the sensors according to equation~\ref{eq:MinimumSensorDistance}, replacing $\boldsymbol{\nu}_c$ with $\boldsymbol{\nu}_{cy}$.

Although such sensors avoid errors due to aliasing, they attenuate the high spatial frequency variations in the observations. Therefore, any procedure applied to estimate the original field or the underlying connectivity structure from these band-limited observations has the potential to underestimate the high spatial frequency components of both the field and the connectivity kernel. This motivates the use of sensors with wider bandwidths (narrower in space). Nevertheless, this choice would require more sensors in order to satisfy Shannon's sampling theorem for a given spatial region. Therefore, a compromise needs to be found between the bandwidth of the sensor response, the accuracy of the estimation results, the number of sensors and the computational demands of the estimation procedure when designing experiments.

@@

The unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{eqnarray}\label{eq:sigmapoints1}
	\mathcal X_{0}&=&\bar x \\
	\mathcal X_{i}&=&\bar x+\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, L \\
	\mathcal X_{i}&=&\bar x-\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_{i- L}, \quad i= L+1, \dots, 2L 
\end{eqnarray}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
\begin{eqnarray}
	\mathbf W_0^{(m)}&=&\frac{\lambda}{ L+\lambda} \\
	\mathbf W_0^{(c)}&=&\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta) \\
	\mathbf W_i^{(m)}&=&\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L, 
\end{eqnarray}
where the superscripts $m$ and $c$ denote mean and covariance and $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$. Since the observation equation is linear (equation~\ref{ObservationEquation}), the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. The cross-covariance matrix of the states, $\mathbf M$, is also required for computing the smoother gain. A summary of the steps in the URTSS algorithm is given in Table~\ref{tab:UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$, and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used.

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a Least Squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new estimate of the state vector sequence. To create the least squares parameter estimator, we first define the $L \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \,\textrm{d}\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from the initialisation or an iteration of the URTSS, we can write
\begin{eqnarray*}
	 \hat{\mathbf x}^b_{1} &=& \mathbf{q}(\hat{\mathbf x}^b_0) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_0+\mathbf e_0 \\
	\hat{\mathbf x}^b_{2} &=& \mathbf{q}(\hat{\mathbf x}^b_1) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_1+\mathbf e_1  \\
	&\vdots& \\
	\hat{\mathbf x}^b_{T}&=& \mathbf{q}(\hat{\mathbf x}^b_{T-1}) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_{T-1}+\mathbf e_{T-1}. 
\end{eqnarray*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathcal W+\mathbf{e}, 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\hat{\mathbf x}^b_{1}\\
		\hat{\mathbf x}^b_{2}\\
		\vdots\\
		\hat{\mathbf x}^b_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\hat{\mathbf x}^b_0)& \hat{\mathbf x}^b_{0}\\
		\mathbf q(\hat{\mathbf x}^b_1)& \hat{\mathbf x}^b_{1}\\
		\vdots & \vdots\\
		\mathbf q(\hat{\mathbf x}^b_{T-1})& \hat{\mathbf x}^b_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathcal W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the LS parameter estimates~\cite{Kailath2000}, $ \hat{\mathcal{W}}$, are
\begin{equation}
	\hat{\mathcal{W}}=(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}

% Results and Discussion can be combined.
\section{Results}\label{ResultsSection} We now provide demonstrations of the estimation procedure using the new methods that are described above. The full neural field model was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Parameters}, unless otherwise explicitly stated for particular experiments. Note that the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is because the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the spatial step size used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field at a single point in time can be seen in Figure~\ref{fig:Figure3}(A). Figure~\ref{fig:Figure3}(B) shows an example of data from the modelled sensors and Figure~\ref{fig:Figure3}(C) shows the power spectral density (PSD) of the simulated time-series. %The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.

>>>>

The mapping between the membrane voltage and the electrophysiological data, denoted by $\mathbf y_t$, is modelled using the observation function that incorporates sensors with a spatial extent by
\begin{equation}
    \label{eq:ObservationEquation}
	\mathbf{y}_t =
	\int_{\Omega}{
	    m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
	\, d\mathbf{r}'} + 
	\boldsymbol{\varepsilon}_t, 
\end{equation}
where $m\left(\mathbf{r}_n-\mathbf{r}'\right)$ is the observation kernel, $\mathbf{r}_n$ defines the location of the electrodes in the field, $n=1,...,n_y$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$ denotes a multivariate normal distribution with mean zero and covariance matrix $\boldsymbol{\Sigma}_{\varepsilon}$.

@@

where $\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}$ denotes the $i^{th}$ column of $\boldsymbol{\Psi}(\mathbf{r}')$ which is a $n_x \times n_{\theta}$ matrix, where $n_x$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity kernel basis functions. This matrix is defined analytically (see the supplementary (\dean{Appendix}) material for analytic convolution of two Gaussians), since it is constant with respect to the states and the unknown parameters. Substituting equation~\ref{eq:DefPsi} into equation~\ref{eq:ReducedForm} gives

@@

Now we define the state disturbance as
\begin{equation}\label{eq:Wt} 
	\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} ) \, d\mathbf{r}},
\end{equation}
which is a zero mean normally distributed white noise process with covariance (see the supplementary material \dean{(Appendix)} for the derivation)
\begin{equation}
	\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top} \, d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
\end{equation}
The observation equation of the reduced model is found by substituting equation~\ref{DefFieldDecomp} into equation~\ref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_t\, d\mathbf{r}'} + \boldsymbol{\varepsilon}_t. 
\end{equation}
The observation equation is linear and can be written in the more compact form
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is 
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}') \, d\mathbf{r}'.
\end{equation}
% \begin{equation}
% 	\mathbf{C} = \left[
% 	\begin{array}{{ccc}} 
% 		c_{1,1} & \dots & c_{1,n_x} \\
% 		\vdots & \ddots & \vdots \\
% 		c_{n_y,1} & \dots & c_{n_y,n_x} 
% 	\end{array}
% 	\right] 
% \end{equation}
% and 
% \begin{equation}
% 	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}') \, d\mathbf{r}'. 
% \end{equation}
Now we have the final form of the state-space model where
\begin{equation}\label{eq:finalformstatespacemodel}
	\mathbf{x}_{t+1} = Q(\mathbf{x}_t) +\mathbf{e}_t
\end{equation}
\begin{equation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t
\end{equation}
and 
\begin{equation}\label{eq:QmatrixForSigmapoints}
	Q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \, d\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t.
\end{equation}

@@

Let the spectral representation of the post-synaptic membrane voltage field at time $t$ be denoted by $V_t(\boldsymbol{\nu})$, where $\boldsymbol\nu$ is the spatial frequency. According to Shannon's sampling theorem, the field must be spatially band-limited for an accurate reconstruction using spatially discrete observations. However, an approximate reconstruction can be obtained if the field is approximately band-limited with 
\begin{equation}
	V_t(\boldsymbol{\nu}) \approx 0 ~ \ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c,
\end{equation}
where $\boldsymbol{\nu}_c$ is a cutoff frequency (typically taken as the -3~dB point). Given such a band-limited field, the distance between adjacent sensors, $\Delta_y$, must satisfy 
\begin{equation}
	\label{eq:MinimumSensorDistance} \Delta_y \leq \frac{1}{2\rho_y\boldsymbol{\nu}_{c}}, 
\end{equation}
where $\rho_y \in \mathbb{R} \ge 1$ is an oversampling parameter. This condition must be satisfied to avoid spatial spectral aliasing effects when reconstructing the hidden dynamic field, $v_t(\mathbf{r})$, using the sampled observations, $\mathbf{y}_t$.

In practice, it is difficult to estimate the bandwidth of the cortex using traditional electrophysiological measurements, possibly preventing placement of sensors in accordance to equation~\ref{eq:MinimumSensorDistance}. However, we envisage it may be possible to estimate the spatial bandwidth using other modalities with higher spatial resolution such as fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Spectral aliasing can be avoided by a proper choice of the spatial sampling distance given the sensors' spectral characteristics. The spatial extent of the sensors results in a spectral low-pass action, thus providing spatial anti-aliasing filtering. Such sensors limit the spatial bandwidth of the observation to a value denoted by $\boldsymbol{\nu}_{cy}$. Aliasing can then be avoided by positioning the sensors according to equation~\ref{eq:MinimumSensorDistance}, replacing $\boldsymbol{\nu}_c$ with $\boldsymbol{\nu}_{cy}$.

Although such sensors avoid errors due to aliasing, they attenuate the high spatial frequency variations in the observations. Therefore, any procedure applied to estimate the original field or the underlying connectivity structure from these band-limited observations has the potential to underestimate the high spatial frequency components of both the field and the connectivity kernel. This motivates the use of sensors with wider bandwidths (narrower in space). This choice would require more sensors in order to satisfy Shannon's sampling theorem for a given spatial region. Therefore, a compromise needs to be found between the bandwidth of the sensor response, the accuracy of the estimation results, the number of sensors and the computational demands of the estimation procedure when designing experiments.

@@

The unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{eqnarray}\label{eq:sigmapoints1}
	\mathcal X_{0}&=&\bar x \\
	\mathcal X_{i}&=&\bar x+\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, n_x \\
	\mathcal X_{i}&=&\bar x-\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_{i- n_x}, \quad i= n_x+1, \dots, 2n_x 
\end{eqnarray}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$ and $n_x$ is the dimension of the state space. The total number of sigma points is $2n_x+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( n_x+\kappa) - n_x, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
\begin{eqnarray}
	\mathbf W_0^{(m)}&=&\frac{\lambda}{ n_x+\lambda} \\
	\mathbf W_0^{(c)}&=&\frac{\lambda}{ n_x+\lambda}+(1-\alpha^2+\beta) \\
	\mathbf W_i^{(m)}&=&\mathbf W_i^{(c)}=\frac{1}{2( n_x+\lambda)} \quad i=1, \dots, 2n_x, 
\end{eqnarray}
where the superscripts $m$ and $c$ denote mean and covariance and $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$. Since the observation equation is linear (equation~\ref{ObservationEquation}), the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. The cross-covariance matrix of the states, $\mathbf M$, is also required for computing the smoother gain, as described in the summary of the steps in the URTSS algorithm is given in Table~\ref{tab:UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$, and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used.

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a Least Squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new estimate of the state vector sequence. To create the least squares parameter estimator, we first define the $n_x \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \,\textrm{d}\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from the initialisation or an iteration of the URTSS, we can write
\begin{eqnarray*}
	 \hat{\mathbf x}^b_{1} &=& \mathbf{q}(\hat{\mathbf x}^b_0) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_0+\mathbf e_0 \\
	\hat{\mathbf x}^b_{2} &=& \mathbf{q}(\hat{\mathbf x}^b_1) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_1+\mathbf e_1  \\
	&\vdots& \\
	\hat{\mathbf x}^b_{T}&=& \mathbf{q}(\hat{\mathbf x}^b_{T-1}) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_{T-1}+\mathbf e_{T-1}. 
\end{eqnarray*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathcal W+\mathbf{e}, 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\hat{\mathbf x}^b_{1}\\
		\hat{\mathbf x}^b_{2}\\
		\vdots\\
		\hat{\mathbf x}^b_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\hat{\mathbf x}^b_0)& \hat{\mathbf x}^b_{0}\\
		\mathbf q(\hat{\mathbf x}^b_1)& \hat{\mathbf x}^b_{1}\\
		\vdots & \vdots\\
		\mathbf q(\hat{\mathbf x}^b_{T-1})& \hat{\mathbf x}^b_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathcal W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the LS parameter estimates~\cite{Kailath2000}, $ \hat{\mathcal{W}}$, are
\begin{equation}
	\hat{\mathcal{W}}=(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}

% Results and Discussion can be combined.
\section{Results}\label{ResultsSection} We now provide demonstrations of the estimation procedure using the new methods that are described above. Data was generated using the neural field model, described in Section~\ref{NeuralModelSection}, to demonstrate the state and parameter estimation framework. All parameters for the model are given in Table~\ref{tab:Parameters}, unless otherwise explicitly stated for particular experiments. Note that the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is because the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the spatial step size used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field at a single point in time can be seen in Figure~\ref{fig:Figure3}(A). Figure~\ref{fig:Figure3}(B) shows an example of data from the modelled sensors and Figure~\ref{fig:Figure3}(C) shows the power spectral density (PSD) of the simulated time-series. %The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.
